---
knit: "bookdown::preview_chapter"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
```

# week4

## 12/04

```bash
stringtie -p 32 -G chrX_data/genes/chrX.gtf -o SRR2980760.gtf –l SRR2980760 SRR2980760.bam
```

```{r}
x <- dir('data-raw/genome-html') %>% plyr::ldply(aves::parse_genome)
x$species

assembly <- sapply(
	dir('data-raw/assembly-html/'),
	. %>% paste0('data-raw/assembly-html/', .) %>% read_html %>% 
		rvest::html_nodes('dd a[href^="/Taxonomy"]') %>% rvest::html_text() %>% 
		stringr::str_extract('\\w+ \\w+')
);
assembly %>% {.[. == 'Anas platyrhynchos']}


read_csv('data-raw/wgs_selector.csv')$prefix_s %>% aves::fetch_TSA()

'GFNW01' %>% paste0('data-raw/TSA-html/', .) %>% read_html() %>% 
	rvest::html_nodes('tr td a[href^="https://www.ncbi.nlm.nih.gov/sra"]') %>%
    	rvest::html_text();
```

differences bewteen gcf and gac

```{r}
mclapply(
	dir('data-raw/assembly-html/'),
	. %>% paste0('data-raw/assembly-html/', .) %>% read_html %>% 
		rvest::html_node('.portlet_content a[href*=GCF][href*=ftp]') %>% rvest::html_attr('href')
) -> gcf

mclapply(
	dir('data-raw/assembly-html/'),
	. %>% paste0('data-raw/assembly-html/', .) %>% read_html %>% 
		rvest::html_node('a[href*=GCA]') %>% rvest::html_attr('href')
) -> gca

gcf %>% unlist %>% na.omit()
```


SRA for Sphenisciformes & Casuariiformes


## 12/05

http://ccb.jhu.edu/software/stringtie/index.shtml?t=manual#output

novel isoforms (i.e. assembled transcripts present in the samples but missing from the reference annotation)

- download OrthoDB

```
http://www.orthodb.org/v9.1/download/odb9v1_metazoa_fasta.tar.gz
http://www.orthodb.org/v9.1/download/odb9v1_genes.tab.gz
http://www.orthodb.org/v9.1/download/odb9v1_levels.tab.gz
http://www.orthodb.org/v9.1/download/odb9v1_species.tab.gz
http://www.orthodb.org/v9.1/download/odb9v1_OGs.tab.gz
http://www.orthodb.org/v9.1/download/odb9v1_OG2genes.tab.gz
http://www.orthodb.org/v9.1/download/README.txt
```

on workstation 
```bash
cat OrthoDB/odb9v1_OGs.tab | grep -P '\t8782' | join OrthoDB/odb9v1_OG2genes.tab - | awk '{print $1 "\t" $2}' > hamstr/OG2genes.tsv
```

assembly statistics

```{r}
mclapply(
	dir('data-raw/assembly-html/'),
	. %>% {
		name <- paste0('data-raw/assembly-html/', .) %>% read_html %>% 
			rvest::html_nodes('dd a[href^="/Taxonomy"]') %>% rvest::html_text() %>% 
			stringr::str_extract('\\w+ \\w+');
		paste0('data-raw/assembly-html/', .) %>% read_html %>% 
			rvest::html_node('table[summary]') %>% rvest::html_table() %>% 
			tidyr::spread(key = X1, value = X2) %>% tibble::as_tibble() %>%
			tibble::add_column(species = name, .before = 1)
	}
) %>% plyr::ldply() -> assembly_stat

tempfile() %T>% write_csv(assembly_stat, .) %>% read_csv



aves_tax_id <- mclapply(
	dir('data-raw/assembly-html/'),
	. %>% paste0('data-raw/assembly-html/', .) %>% read_html %>% 
			rvest::html_nodes('dd a[href^="/Taxonomy"]') %>% rvest::html_attr('href') %>% str_extract('\\d+')
) %>% plyr::laply(as.character) 

read_tsv('data-raw/odb9v1_species.tab', F, 'cciic') %>% dplyr::filter(X1 %in% aves_tax_id);
```

open http://www.orthodb.org/, check aves, unfold all (aves + 4 order), save all

```{r}
aves_name <- read_html('data-raw/Ortholog Search _.html') %>% 
	rvest::html_nodes('ul[style*=block]') %>% 
	rvest::html_nodes('span.tree-name:not(.tree-folder)') %>% 
	rvest::html_text() %>% str_replace_all('[\n\t]', '');

orthodb <- read_tsv('data-raw/odb9v1_species.tab', c('id', 'species', 'gene', 'OG', 'type'), 'cciic') %>% 
	dplyr::filter(species %in% aves_name) %>% aves::bind_taxonomy()

```


Finally, I decide to choose 9031 'Gallus gallus', 59729 'Taeniopygia guttata' & 128390 'Nipponia nippon'

For chicken, Zebra finch, and one Pelecaniformes species to represent core waterbirds

## 12/06

```{r}
omics %>% plyr::ddply('species', . %>% {
	result <- .;
	
	if ('protein' %in% .$source) {
		result = dplyr::filter(., source == 'protein');
	} else if ('TSA' %in% .$source) {
		result = dplyr::filter(., source == 'TSA');
	} else if ('SRA' %in% .$source) {
		result = dplyr::filter(., source == 'SRA');
	}
	
	dplyr::slice(result, 1);
	
}) %>% {
	order_3 <- table(.$order) %>% 
	dplyr::as_tibble() %>% dplyr::filter(n > 2) %>% {.[[1]]};
	dplyr::filter(., order %in% order_3) %>% dplyr::as_tibble()
} %>% dplyr::filter(source == 'SRA') %>% {.$id} -> Run

#" do not use fetch since downloaded html file doesn't contains the form
#paste0('https://www.ebi.ac.uk/ena/data/view/', Run) %>% 
#    plyr::l_ply(browseURL)

run <- yaml::yaml.load_file('data-raw/sra.yaml')

setequal(
plyr::laply(run, . %>% str_extract('\\wRR\\d+(?=[\\._])') %>% unique),
Run)
 
setequal(names(run), Run)

yaml::yaml.load_file('data-raw/sra.yaml') %>% unlist() %>% 
	str_replace('([\\w_\\.]+)(\\.gz)$', '\\1.gz\n\tout=\\1') %>% 
	write_lines('data-raw/fastq.md')
```


## 12/07

> no matter you prefer genome or SRA, the workload is almost the same.

```{r}
#" favor genome
aves::omics %>% plyr::ddply('species', . %>% {
	result <- .;
	
	if ('protein' %in% .$source) {
		result = dplyr::filter(., source == 'protein');
	} else if ('TSA' %in% .$source) {
		result = dplyr::filter(., source == 'TSA');
	} else if ('genome' %in% .$source) {
		result = dplyr::filter(., source == 'genome');
	}
	
	dplyr::slice(result, 1);
	
}) %>% {table(.$source)}

#" favor SRA
aves::omics %>% plyr::ddply('species', . %>% {
	result <- .;
	
	if ('protein' %in% .$source) {
		result = dplyr::filter(., source == 'protein');
	} else if ('TSA' %in% .$source) {
		result = dplyr::filter(., source == 'TSA');
	} else if ('SRA' %in% .$source) {
		result = dplyr::filter(., source == 'SRA');
	}
	
	dplyr::slice(result, 1);
	
}) %>% {table(.$source)}
```

in hamstr, `blosum62.bla` should be `BLOSUM62.bla`

> ublast is much faster than blastp (3.5x)

```
hamstr -sequence_file=testset-prot.fa -taxon=test2 -hmmset=modelorganisms_hmmer3 -refspec=DROME -central -force -cpu 32
blastp
real    6m56.809s
user    173m58.157s
sys     7m42.249s

ublast
real    1m57.670s
user    29m45.085s
sys     13m57.842s
```

```
[entomo@localhost output]$ time ../software/hamstr.v13.2.6/bin/hamstr -protein -sequence_file=GCF_000002315.4_Gallus_gallus-5.0_protein.faa -taxon=Gallus_gallus -hmmset=aves -refspec=Gallus_gallus -blastpath=/path/to/data/aves/hamstr/blast_dir -hmmpath=/path/to/data/aves/hamstr -representative -central -force -cpu 32 -ublast &> /dev/null;

real    147m3.959s
user    3270m35.124s
sys     1163m17.227s
[entomo@localhost output]$ time ../software/hamstr.v13.2.6/bin/hamstr -protein -sequence_file=GC0_000002315.4_Gallus_gallus-5.0_protein.faa -taxon=Gallus_gallus -hmmset=aves -refspec=Gallus_gallus -blastpath=/path/to/data/aves/hamstr/blast_dir -hmmpath=/path/to/data/aves/hamstr -representative -central -force -cpu 32 -ublast -hit_limit 10 &> /dev/null

real    32m5.598s
user    898m51.634s
sys     105m7.312s
```

## 12/08

refercence species (abandoned) 
```r
protein <- aves::genome %>% dplyr::filter(protein != '') %>% aves::bind_taxonomy() %>% 
    dplyr::mutate(species = str_replace(species, ' ', '_'));

protein %>% dplyr::mutate(input = basename(protein)) %>% dplyr::select(species, input) %>% 
	plyr::d_ply('species', . %>% {
		fasta <- bioinfor::read_fasta(paste0('raw/protein/', .$input));
        names(fasta) %<>% str_replace('\\.[\\w\\W]+$', '');
        bioinfor::write_fasta(fasta, paste0('protein/', .$species, '.faa'));
	}) 

ref_species <- character(nrow(protein));
ref_species[protein$order %in% aves::stratum$order[aves::stratum$Galloanseres]] = 'Gallus_gallus';
ref_species[protein$order %in% aves::stratum$order[aves::stratum$Telluraves]] = 'Taeniopygia_guttata';
ref_species[protein$order %in% aves::stratum$order[aves::stratum$`Aequornithia+`]] = 'Nipponia_nippon';
ref_species[protein$order %in% aves::stratum$order[aves::stratum$`Other Neoaves`]] = 'Taeniopygia_guttata,Nipponia_nippon -relaxed';
ref_species[ref_species == ''] = 'Gallus_gallus';
```


> NCBI genome annotation is better than Jarvis 2014 so we use former

```{r}
aves::genome %>% dplyr::filter(protein != '') %>% 
	dplyr::mutate(file_name = basename(protein)) %>%
	dplyr::mutate(species = str_replace(species, ' ', '_')) %>%
	dplyr::select(species, file_name) %>% 
	plyr::d_ply('species', . %>% {
		fasta <- read_lines(paste0('raw/protein/', .$file_name[1]));
        names(fasta) %<>% str_replace('\\.[\\w\\W]+$', '');
        write_lines(fasta, paste0('protein/', .$species[1], '.faa'));
	})

dir('bird_phylogenomics_data/gene_and_repeat_annotation/coding_sequences/pep/') %>% 
    plyr::l_ply(. %>% {
        fasta <- read_lines(paste0('bird_phylogenomics_data/gene_and_repeat_annotation/coding_sequences/pep/', .));
        names(fasta) %<>% str_replace(' [\\w\\W]*$', '');
        write_lines(fasta, paste0('protein/', str_replace(., '\\.gz', '')));
    })


dir('protein', '\\.pep', full.names = T) %>%  {c(., str_replace(., 'pep', 'faa'))} %>% 
    file.info %>% tibble::rownames_to_column() %>% dplyr::select(1:2) %>% 
    dplyr::mutate(species = str_extract(rowname, '(?<=/)[\\w_]+'), size = size/1e6) %>% 
    dplyr::mutate(type = str_extract(rowname, '[\\w]+$')) %>% dplyr::select(2:4) %>%
    ggplot2::ggplot() + ggplot2::geom_point(ggplot2::aes(species, size, color = type)) + ggplot2::coord_flip() + ggplot2::scale_color_discrete(labels = c('NCBI', 'Jarvis')) + ggplot2::labs(y = 'size (MB)', title = 'NCBI genome annotation is better than Jarvis 2014') + ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

## 12/09

analyse hmmer hits number

```r
OG <- dir('hamstr/aves/hmm_dir/')
species <- str_extract(dir('protein', 'hmm_'), '(?<=search_)\\w+(?=_aves)') %>% str_replace('_', ' ');
hmm_hit <- dir('protein', 'hmm_', full.names = T) %>% plyr::llply(. %>% paste0('/', OG) %>% mclapply(. %>% read_file %>% str_count('>>')) %>% unlist) 


seq_along(hmm_hit) %>% plyr::ldply(. %>% {
    tibble::data_frame(
        species = species[.], OG = OG, hmm_hit = hmm_hit[[.]])
}) %>% tibble::as.tibble()

write_csv(.Last.value, 'protein_hmm_hit.csv')
```

## 12/10

transcript assembly

1. Bridger(2015)

https://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0596-2
https://sourceforge.net/projects/rnaseqassembly/files/?source=navbar

1. SOAPdenovo-Trans(2014)
https://academic.oup.com/bioinformatics/article/30/12/1660/380938
https://github.com/aquaskyline/SOAPdenovo-Trans

1. Oases(2012)

https://academic.oup.com/bioinformatics/article/28/8/1086/195757
https://github.com/dzerbino/oases

1. Trinity(2011)

https://www.nature.com/articles/nbt.1883
https://github.com/trinityrnaseq/trinityrnaseq/wiki

1. Trans-ABySS(2010)

https://www.nature.com/articles/nmeth.1517
https://github.com/bcgsc/transabyss


others

1. Optimizing for phylogenomics (last 4)
https://bmcgenomics.biomedcentra## 12/04












































